\CRANpkg{lg}: An R package for Local Gaussian Approximations

### 2021-02-16 response/response.Rmd

> *Overall evaluation*
>
> *Measuring and visualizing the structure(s) of temporal dependencies and dependencies among variables – or aspects thereof – is highly relevant in all areas and sub-disciplines of statistics (e.g. biometry, econometrics etc.). Therefore, providing a software package that implements the local Gaussian correlation, a measure of linear and non-linear dependencies recently proposed by Tjøstheim and Hufthammer (2013) is very useful since it allows a broader audience to apply this concept in empirical research. In this regard presenting such a software package in a scientific journal seems to further support applied statisticians in using this new concept of dependence. However, the author fails to convince me that the manuscript constitutes a serious scientific piece at this stage which deserves publication in a well respected scientific journal. A number of reasons have lead to this conclusion.*

**My response:** The reviewer is exactly right. There is a lot of interest in the measurement, visualization and modeling of statistical dependence in many empirical disciplines. The purpose of this paper is to make recent developments within local Gaussian approximations that have appeared in the statistical literature in recent years available to a broader audience of practitioners. I appreciate the comments by the reviewer below, and I hope that the revised paper satisfies the high scientific standards of *The R Journal*.

> *To exemplify this, the section on ‘Statistical background’ reads as if compiled from different sources; In particular, the paragraphs are not well connected in my view.*

**My response:** The purpose of this section is to give the reader a brief overview of the statistical methods that have been implemented in the **lg**-package, without going too much in to the technical details. I have revised this section carefully, with particular emphasis on improving the transitions between the paragraphs.

@@

>  *Further the explanation of the basic idea of the GLC measure in the section ‘Statistical background’ could be improved, thereby making the concept accessible for a broader audience and making this paper more ‘self-contained’.*

**My response:** Thank you for this input. The reviewer is correct to point out that the section on "Statistical background" is fairly short and does not go into great detail. The composition of this section is the result of a careful trade-off. On the one hand I fully agree that a more comprehensive section would be desireable, because that would allow the reader to appreciate the technical nuances within the topic of the LGC. On the other hand, it is the purpose of this paper to demonstrate the practical implementation of a large and diverse set of statistical methods. It is my conclusion that this section should not be much longer than it is in its current form (approx. 4 pages), which naturally puts a limit on the level of detail. I have instead opted to provide a rather acessible birds-eye view of the relevant literature, which I believe is useful to the readership of *The R Journal*. This section has been heavily revised as prompted by the reviewer in his/her comment above..

@@

> *Also, the whole manuscript is not well-structured: The section with the illustration of the lg package is called ‘The lg-object’ and more importantly is confusing since, for instance, all three different estimation methods implemented are outlined, although not necessary for the (empirical) example chosen by the author;*

**My reponse:** The paper is structured in the same way as the package it describes. There are two steps to every analysis; in the first step the practitioner must make several modeling choices which to a large degree is guided by the respective references. This step is in common for all applications. I have created a new title for this section to clarify the structure; it is now called "The first step: Creating the lg-object". The various applications are then described separately and organized under informative headlines.

I am not entirely sure how to address the reviewer's second remark. The examples in this paper are chosen to illustrate key steps in the implementation of the methods. It is not my intention to demonstrate all possible combinations of the modeling choices in examples, as that would likely reduce the value of the paper as a general, readable and fairly brief introduction to the relevant R-functions.

@@

> *A section with a summary or concluding remarks is fully missing!*

**My response:** I have added some conluding remarks in the end of the paper.

@@

> *Furthermore, the code chunks presented in the paper could be augmented by comments explaining what the different lines of code do; For example, in the section ‘Graphics’ it is not clear what expand.grid does.*

**My response:** Thank you for pointing this out. Several of the somewhat longer code chunks were not properly commented. Has been fixed.

@@

> *In all, this version of the manuscript, unfortunately, appears to be incomplete and as if submitted under time pressure.*

**My response:** I can ensure the reviewer and the editor that this is not the case. The manuscript has been prepared carefully, under no time pressure whatsoever. I do thank the reviewer for several useful comments, which I believe has contributed to a significant improvement in the presentation of the material.

Here are my responses to the specific issues addressed by the reviewer:

> *1. As far as I see is your manuscript missing a final section that summarizes the main ideas (e.g. functions here) from the paper and provides some concluding remarks and perhaps an outlook (e.g. what will be added to package in the future). I strongly recommend to provide such a section.*

**My response:** Has been added.

> *2. Are you sure that Equation (1) is correct? I think $x$ should be $x_0$ . Further I do not understand $y$ and assume $y$ should be $X_i$?*

**My response:** Yes, this equation is correct. The $x$ is the location of the local estimate as defined in the discussion leading up to equation (1), and $y$ is the integration variable in the penalty term of the local likelihood function. This notation is consistent with the literature on local likelihood estimation. The second term does in fact not depend directly on the observations, and the reviewer is quite right to point out that this is not obvious. I have therefore added a comment about this, as well as a reference to a much more detailed justification of the local likelihood function.

> *3. Are you sure that ‘the same test statistic and bootstrap procedure as for independence’ can be used with time series data? Is not block-bootstrap necessary in this case? Please explain briefly! (p. 3)*

**My response:** In this particular case we can use the standard bootstrap, because serial dependence in a time series can be investigated by considering the statistical dependence between $X_t$ and $X_{t-k}$ directly. This is a parallell to the autocorrelation function, which for a stationary time series is defined as the ordinary Pearson correlation $\gamma_k = Corr(X_t, X_{t-k})$. I have changed the formulation here so that it directly states that we obtain critical values by permutations. The block bootstrap is then introduced in the next paragraph, where it is needed to test for serial cross-dependence.

> *4. I understand that the bandwidth controls how many observations are used to compute the correlation coefficient of one grid. Does lg also allow to extract the (exact) number of observations used to compute the different correlations coefficients (as depicted e.g. in Fig. 1)? I feel that from the perspective of an applied statistician, this figure could provide an intuition about how reliable one or the other coefficient is.*

**My response:** The bandwidth does not define a finite region over which the usual correlation is calculated; this is, it is not the conditional correlation, and the bandwidth does not correspond to an exact number of observations. Figure 1 shows the estimated local Gaussian correlation plotted out on a grid chosen by the user, and I have pointed this out in the first line of the paragraph that introduces this figure.

Although one can discuss the "effective number of observations" in the classical nonparametric sense, this has not been done in the literature on the LGC, and hence not a consept in the *lg*-package.

> *5. Is it possible to test for significant differences in the correlation coefficients, for instance, from the lower left, middle and upper right in Fig. 1?*

**My response:** That may certainly be possible. To the best of my knowledge, however, this is not a test that has been introduced in the literature. It is therefore not implemented in the **lg**-package.

> *6. Page 1: I suggest to change the statement ‘clever and creative techniques’ by a more scientific formulation.*

**My response:** This statement has been taken out. Thank you.

### 2021-02-16 otneim.tex
First pass of revision

\citet{hjor:jone:1996} provide one such idea. They consider a parametric family $F_{\theta}$, but instead of searching for a single parameter value $\theta_0$, for which $f_X(x) = f(x;\theta_0)$ (or approximately so), they rather assert that different members of $F_{\theta}$ may approximate $f_X$ locally in different parts of its domain. In other words, they seek to estimate a parameter \emph{function} $\theta_0(x)$ for which $f_X(x) = f(x;\theta_0(x))$ (or approximately so), and do this by maximizing a \emph{local likelihood function} in each point $x$:
\begin{align}
\widehat\theta(x) 
&= \arg\max_{\theta \in \Theta} L_n(\theta, x) \nonumber\\ 
&= \arg\max_{\theta \in \Theta} \frac{1}{nb}\sum_{i=1}^n K\left(\frac{X_i - x}{b}\right)\log f(X_i; \theta) - \int \frac{1}{b}K\left(\frac{y - x}{b}\right)f(y;\theta)\,\textrm{d}y,
\label{eq:loclik}
\end{align}
where, again, $K$ is a symmetric density function and $b$ is a bandwidth parameter that controls the smoothness of the estimate.

@@

In Figure \ref{fig:initial-example} we have plotted the estimated LGC for two bivariate data sets; 1000 simulated observations from a binormal distribution having correlation equal to 0.5, and the daily return on the CAC40 and FTSE100 stock indices on 1000 consecutive trading days starting on May 5th 2014. In the first panel, we see that the estimated local correlation coincides with the global correlation, except for the estimation error. In the second panel we see clearly that the local correlation, and thus the dependence, is stronger in the lower left and upper right regions of the distribution than in central parts. The phenomenon of local dependence is well known in the financial literature, and using the LGC it can be measured, interpreted, and visualized in a natural way. The interpretation of this particular figure is that extreme observations on the two stock indices are more strongly dependent than the less extreme observations.

@@

\section{The \texttt{lg}-object} 
\label{chap:lgobject}

@@

\begin{example}
R> lg_object <- lg_main(x,
R+                      est_method = "5par_marginals_fixed",
R+                      transform_to_marginal_normality = TRUE,
R+                      bw_method = "cv")
\end{example}

@@

\begin{example}
R> lg_object$bw

@@

\begin{example}
R> lg_object <- lg_main(x,
R+                      est_method = "5par_marginals_fixed",
R+                      transform_to_marginal_normality = TRUE,
R+                      bw_method = "plugin")
R> lg_object$bw

@@

\begin{example}
R> lg_object <- lg_main(test_x, 
R+                      est_method = "5par",
R+                      transform_to_marginal_normality = TRUE)
R> test_result <- ind_test(lg_object, n_rep = 100)
R> test_result$p_value

@@

\begin{example}
R> grid <- expand.grid(seq(-3, 3, length.out = 7),
R+                     seq(-3, 3, length.out = 7))
R> 
R> x <- mvtnorm::rmvnorm(500, sigma = matrix(c(1, rho, rho, 1), 2))
R> lg_object <- lg_main(x, 
R+                      est_method = "5par", 
R+                      transform_to_marginal_normality = FALSE,
R+                      plugin_constant_joint = 4)
R> dlg_object <- dlg(lg_object, grid = grid)
R> 
R> # Make a dependence map using default setup
R> corplot(dlg_object1)
\end{example}

@@

>>>>

\citet{hjor:jone:1996} provide one such idea. They consider a parametric family $F_{\theta}$, but instead of searching for a single parameter value $\theta_0$, for which $f_X(x) = f(x;\theta_0)$ (or approximately so), they rather assert that different members of $F_{\theta}$ may approximate $f_X$ locally in different parts of its domain. In other words, they seek to estimate a parameter \emph{function} $\theta_0(x)$ for which $f_X(x) = f(x;\theta_0(x))$ (or approximately so), and do this by maximizing a \emph{local likelihood function} in each point $x$:
\begin{align}
\widehat\theta(x) 
&= \arg\max_{\theta \in \Theta} L_n(\theta, x) \nonumber\\ 
&= \arg\max_{\theta \in \Theta} \frac{1}{nb}\sum_{i=1}^n K\left(\frac{X_i - x}{b}\right)\log f(X_i; \theta) - \int \frac{1}{b}K\left(\frac{y - x}{b}\right)f(y;\theta)\,\textrm{d}y,
\label{eq:loclik}
\end{align}
where, again, $K$ is a symmetric density function and $b$ is a bandwidth parameter that controls the smoothness of the estimate. The second term in the local likelihood function is a penalty that ensures that the estimated density $\widehat f_X(x) = f(x; \widehat\theta(x))$ converges correctly to the true density function $f_X(x)$ as the sample size $n$ increases to infinity and the bandwidth $b$ decreases towards zero. See \citet{hjor:jone:1996} for a detailed discussion about this construction.

@@

In Figure \ref{fig:initial-example} we have plotted the estimated LGC for two bivariate data sets on a grid; 1000 simulated observations from a binormal distribution having correlation equal to 0.5, and the daily return on the CAC40 and FTSE100 stock indices on 1000 consecutive trading days starting on May 5th 2014. In the first panel, we see that the estimated local correlation coincides with the global correlation, except for the estimation error. In the second panel we see clearly that the local correlation, and thus the dependence, is stronger in the lower left and upper right regions of the distribution than in central parts. The phenomenon of local dependence is well known in the financial literature, and using the LGC it can be measured, interpreted, and visualized in a natural way. The interpretation of this particular figure is that extreme observations on the two stock indices are more strongly dependent than the less extreme observations.

@@

\section{The first step: Creating the \texttt{lg}-object} 
\label{chap:lgobject}

@@

\begin{example}
R> # Create the lg-object with bandwidths chosen by cross-validation
R> lg_object <- lg_main(x,
R+                      est_method = "5par_marginals_fixed",
R+                      transform_to_marginal_normality = TRUE,
R+                      bw_method = "cv")
\end{example}

@@

\begin{example}
R> # Print out the bandwidths
R> lg_object$bw

@@

\begin{example}
R> # Make the lg-object with plugin bandwidths
R> lg_object <- lg_main(x,
R+                      est_method = "5par_marginals_fixed",
R+                      transform_to_marginal_normality = TRUE,
R+                      bw_method = "plugin")
R> # Print out the bandwidths
R> lg_object$bw

@@

\begin{example}
R> # Make the lg-object
R> lg_object <- lg_main(test_x, 
R+                      est_method = "5par",
R+                      transform_to_marginal_normality = TRUE)
R> # Perform the independence test
R> test_result <- ind_test(lg_object, n_rep = 100)
R> # Print out the p-value of the test
R> test_result$p_value

@@

\begin{example}
R> # Make a regular grid in the domain of the distribution
R> grid <- expand.grid(seq(-3, 3, length.out = 7),
R+                     seq(-3, 3, length.out = 7))
R> 
R> x <- mvtnorm::rmvnorm(500, sigma = matrix(c(1, rho, rho, 1), 2))
R> lg_object <- lg_main(x, 
R+                      est_method = "5par", 
R+                      transform_to_marginal_normality = FALSE,
R+                      plugin_constant_joint = 4)
R> dlg_object <- dlg(lg_object, grid = grid)
R> 
R> # Make a dependence map using default setup
R> corplot(dlg_object1)
\end{example}

@@

\section{Conclusion}

The statistical literature has seen a number of applications of local Gaussian approximations in the last decade, covering several topics in dependence modeling and inference, as well as the estimation of multivariate density and conditional density functions. In this paper, we demonstrate the implementation of these methods i the R programming language using the \CRANpkg{lg}-package, as well as the graphical representation of the estimated local Gaussian correlation. The package is complete, in the sense that all major methods that have been published within this framework is now easily accessible to the practitioner. The package is also modular; Future methodological developments using local Gaussian approximations may be added to the package with little effort.
